---
title: "Working with Data from a BioRad CFX 96-well Instrument"
author: "Alexandra Emmons"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BioRad_CFX_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(qpcrplotR)

```
## Introduction
This vignette documents how data from a BioRad CFX 96 well instrument may be analyzed by the R package *qpcrplotR*. The purpose of this vignette is to demonstrate how to load, normalize, and plot data outputted from a BioRad CFX 96 well qPCR machine.

Most of this information may also be found throughout the R documentation. 

## Loading data

The *qpcrplotR* package manages data outputted from three different qPCR machines: 1) the BioRad CFX 96-well instrument, the quantamaster via qPCR machine, and the BioRad2 qPCR instrument. *qpcrplotR* takes data from each of these instruments and standardizes them to a common format, so that the data can be normalized and visualized using a variety of available functions. 

This vignette will focus on data from a BioRad CFX 96-well qPCR instrument. The load function for this type of data is `load_bioradcfx()`. The inputed data must be a comma delimited file. The BiorRAd CFX exports files as .xlsx files, so make sure the data is converted to a .csv file prior to using the *qpcrplotR* package.

Here is an example of how data is imported with this function:

```{r}
library(dplyr)
library(tidyverse)
quant <- load_bioradcfx("../data/test.csv")

```

Using the `load_bioradcfx()` function, sample names, measured copy number, means, and standard deviations are all imported and converted into a tibble table to increase readability.To perform statistical analysis on your data it is imperative that you use at least three replicates per sample.

After using the `load_bioradcfx()` function the data should look similar to the following:

```{r}
head(quant)

```
This data represents bacterial 16S gene abundances that were coextracted from human bone samples. The quantity of each sample is determined based on a standard curve, denoted as samples S1-S5. 


## Data quality control and normalization

Once data has been loaded using the load function, there are a variety of different functions that may be used for data quality control and normalization of data. Here is a table of these functions with some brief descriptions:

| Function     | Description |
|:------------:|:------------|
|`repsout()`   | Removes outliers from dataset |
|`norm_pd()`   | Normalizes data by Pipette volume and dilution factor |
|`norm_sd()`   | Normalizes a subset of samples by the dilution factor and the entire set by pipette volume |
|`norm_biom()` | Normalizes data by extraction elution volume and biomass input |

For our BioRad CFX data we will use the `repsout()` and `norm_pd()` functions so that we can first remove any outliers from the data and then normalize the data by inputted pipette volume, the dilution factor, and biomass input. 

Removing outliers using `repsout()` requires the dataframe created from our original load function. Outliers are removed using this function based on any inputed number of standard deviations, which is the second argument required. For example, if the second argumnet is set to 2, then only values within two standard deviations of the mean will be included in the resulting data frame, giving the user control over outlier removal.

The `norm_pd()` function will then use the resulting dataframe as its first argument. The two other arguments are user-input depending on how the qPCR assay was run. These two arguments are the pipetted volume of sample and the dilution factor. The example qPCR data in this vignette used 2-$\mu$L of sample and was diluted 10-fold. Therefore we will use the values 2 and 10 for these arguments. We do this as follows:

```{r}
quant1 <- repsout(quant, 2)
quant2 <- norm_pd(quant1, 2, 10)
```

If for some reason you did not dilute all samples prior to qPCR, the function `norm_sd()` can be used to normalize only a subset of your samples by the dilution factor. In order for `norm_sd()` to work, a vector of the sample names that were diluted must be created:
```{r}
x<-c('1101-0001|1101-0002|1101-0003')
```

This vector will be used as an argument in the function to designate which samples should be normalized by the dilution factor.As you can see the sample names in the example subset are 1101-0001, 1101-0002, and 1101-0003.This function requires an inputed data frame (quant1) that is the result of `repsout()`, a vector (x) containing sample information, and the dilution factor (10). An example is as follows:
```{r}
example<-norm_sd(quant1, x, 10)
```

After either `norm_pd()` or `norm_sd()` is applied to your data, you can then use the `norm_biom()` funtion to normalize your data by a biomass input.  For example, if DNA was isolated from 0.25g of bone powder, you may want to determine the abundance of your targeted gene per gram of bone powder. You can use the `norm_biom` function to do this. This function requires the following arguments: a data frame resulting from either `norm_pd()` or `norm_sd()`, the volume of your final extraction (elution) volume, and the mass of the biomass input. An example is as follows:
```{r}
quant3<-norm_biom(quant2, 100, 0.25, "Mean.Copies/g")
```

Finally, metadata may be associated with your samples prior to plotting using the `meta_merge()` function. 

```{r}
quant4<-meta_merge("../data/test_metadata.csv", quant3)
head(quant4)
```
As you can see, the resulting data frame (quant4) now includes additional information including the Agency.Id, referring to the biological replicate ID, the Body.region that was sampled, and the Human.DNA quantity associated with that region. 


ALTERNATIVELY, `meta_merge` may be used following `norm_pd` or `norm_sd` and the function `norm_biom2`may instead be used to normalize your data by biomass input. This function is useful if the initial biomass input differs by samples, meaning exact masses were taken for each sample prior to extraction. 
```{r}
quant5<-meta_merge("../data/test_metadata.csv", quant2)
norm_biom(quant5, 100, "Mean.Copies.gdw", "Soil.dw.")
```

## Plotting data

Once data has been checked for quality and normalized, the last step is to look at the data. *qpcrplotR* can't interpret your results for you; however, it will provide a preliminary plot of your data, so that you can guage how samples differ and whether certain samples should be requantified. To do this, use the `plot_pcr()` function. It plots your data using the R package *ggplot2* so that you may begin visualizing your data.

Now that the data is normalized it can be plotted as follows:

```{r}
library(ggplot2)
plot1 <- plot_pcr(quant4, "Sample", "Mean.Copies/g", "Body.region")
print(plot1)

```

To see how these functions can be applied to other types of qPCR data, checkout more of the vignettes associated with *qpcrplotR*.
